### 2022-06-29 paper/paper.tex
Paper cleanup

\documentclass[12pt,a4paper]{article}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{booktabs}
\usepackage[latin2]{inputenc}
\usepackage{t1enc}
\usepackage[mathscr]{eucal}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{graphics}
\numberwithin{equation}{section}
\usepackage{hyperref}
\usepackage[margin=2.9cm]{geometry}
\usepackage[outdir=./images/]{epstopdf}
\epstopdfsetup{outdir=./images/}
\usepackage{algorithm}
\usepackage{algpseudocode}

\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}

\def\numset#1{{\\mathbb #1}}

\begin{document}

\title{Estimating Gaze Duration Error from Eye Tracking Data}

\author{
John Hawkins \\ email \href{mailto:john.hawkins@playgroundxyz.com}{john.hawkins@playgroundxyz.com} \\
}

\maketitle

\section{Abstract}

Eye tracking applications produce a series of gaze fixation points that can be attributed to
objects within a subject's field of vision. Error is typically measured on the basis of individual
gaze fixation point measurements. These applications are often used to infer a gaze duration
metric from a series of fixation measurements. There is no direct method for infering the error in
a gaze duration measurement from an error in fixation points.

In this work we develop an algorithm for estimating
the error bounds on gaze duration measurement through monte carlo simulation based on the content
of an eye tracking calibration log file. We provide this algorithm as an open source application
that allows for robust estimates of the error distribution of gaze duration measurements. 
We use this application to conduct experiments on the expected error bounds for different duration
measurements across a range of session lengths and area of interest positions within the viewport
of mobile devices. The results indicate that error in gaze duration estimation is sensitive to
the size of fixation error as well as the proportion of total field of view that the object occupies
and the proportion of viewing time that was taken with gaze fixation.

\section{Introduction}

Eye tracking is an important technology with a wide variety of applications. It permits evaluation
of software user interfaces\cite{Harezlak2015}, the development of new forms of software interaction, as well as a potential method of biometric identification\cite{Kasprowski2018}. 
Eye tracking has become a core tool for empirical investigations 
into human behaviour and is widely used as a method of measuring explicit attention to visual stimuli. 
It has been applied to study psychological phenomena ranging from cognition\cite{Brunye2019}
to mental health\cite{Duque2014,Rudich-Strassler2022}, and is now routinely used to evaluate 
advertising \cite{Hervet2011}. Eye tracking technology has allowed marketing researchers 
to study many factors that contribute to effective advertising, 
including brand recall \cite{Wedel2000},
the capture and tranfer of attention \cite{Pieters2004}, 
the impact of images of faces \cite{Djamasbi2010},
the attention effects of animation \cite{Hamborg2012},
and the relationship with social media posts \cite{Barreto2013}.

Increasingly, eye tracking applications are built using machine learning models trained on 
data sets containing combinations of facial images and fixation coordinates collected 
through careful application design. The resulting gaze fixation prediction model can 
then be applied sequentially to determine how long a subject fixated on a given object.
Thus, gaze duration is derived from aggregation of gaze fixation.

Gaze fixation models will typically have an error profile that is dependent on both 
the nature of the media being presented, the variety of faces and lighting conditions 
in the training data. In addition, it has been observed that the error profile varies 
depending on the location of the true fixation point within the subjects field of view.
The errors in fixation point measurement for a given eye tracking solution are routinely
adjusted through a process called calibration. Improvements in the callibration process
are an ongoing focus for the development of algorithms\cite{Zhang2014,Hassoumi2019} 
and software tools\cite{ETCAL2018}

When fixation models are used sequentially for the purpose of estimating gaze duration
(as is commonly done in point of salience or advertising media studies), 
then there is the potential for the error
to either cancel out (reducing error in gaze duration estimation) or to compound
(increasing error in gaze duration). Which of these outcomes occurs will depend on the
specifics of the model and the conditions under which it is being used.
Influencing factors include the ratio of measured gaze duration to the total viewing time
and the error profile of the machine learning fixation coordinate model.

The errors of eye tracking data technology can be decomposed into a range of
independent sources that researchers need to consider \cite{Holmqvist2012}.
Unless they are properly addressed these sources of error can manifest themselves
as systematic biases across undesirable dimensions such as the age\cite{Dalrymple2018}
or ethnicity of subjects\cite{Blignaut2013}.

The use of eye fixation models to estimate gaze duration (or dwell time) in Area of
Interest studies (AOI) was discussed by Holmqvist et al. \cite{Holmqvist2012}. The
authors simulated the impact of gaze fixation error by adding noise according to
manufacturer specifications to data sets of low margin areas of interest. We extend 
this idea to calculate probabilistic bounds on the error in gaze duration using
calibration data as a source of noise distribution that is specific to both the
study participants and the device/environment of the study. The result is an open
source application that may be used by a wide variety of reserachers to provide
error bounds on any gaze duration measurements.

\section{Methodology}

We estimate the gaze duration measurement error through a Monte Carlo simulation
using the eye tracking model's calibration file of fixation errors.
The error profile of the eye tracking model allows us to generate sequences of true
fixations and simulate measured fixations with the observed error of the model.
These sequences of true and measured fixation are then converted into a distribution
of expected gaze duration on a specifc area of interest, for a given measurement.

The algorithm consists of two core steps, first estimating the distrubtion of measurements
for a given true duration. Secondly, inverting this into the distribution of true
gaze durtions for a given measurement. Note, that the algorithm requires that the area of
interest be defined as a fixed bounding box and that the distributions are estimated for
a fixed session length. This session length is the total period of time that the area of
interest was in a subject's field of view, and thus represents the maximum possible gaze
duration.

\subsection{Measurement Distribution}

In the first stage of the methodology we provide an estimate of the distribution of measured
durations for all possible true gaze durations. These distributions are produced by generating
random fixation paths across the available screen dimensions, and adding noise that is 
drawn from the calibration data to make it consistent with the observed properties of the gaze model.
This produces a set of distributions that capture the expected
variation in measurement depending on the length of true fixation.

Formally stated this is the probability distribution over measured durations $\hat{d}$ given a
known duration $d$, the session length $s$ and the gaze target location $l$. 
We express this distribution as $P(\hat{d}|d,l,s)$. The
procedure for making this estimation involves iteratively generating random samples
of gaze fixation paths $\mathcal{F}$ consistent with the parameters $d$, $s$ and $l$. 
For each point $f$ in $\mathcal{F}$ we determine a measurement point $\hat{f}$ 
by drawing samples from the eye tracking calibration set. 
The samples from the calibration data are used to determine the $\delta_x$ and $\delta_y$ 
error in the measurement of gaze fixation point $f$. 
We draw these delta values from the calibration file such the
the probability of the point being chosen is propotional to its distance from the point $f$
in the generated path. The complete algorithm
for estimating $P(\hat{d}|d,l,s)$ is shown in Algorithm \ref{alg:p_of_dhat}

\begin{algorithm}
\caption{Estimation of $P(\hat{d}|d,l,s)$}\label{alg:p_of_dhat}
\begin{algorithmic}
  \Require $l,s,C$
  \Ensure $P(\hat{d}|d,l,s)$
  \State $N \gets samples$              \Comment{Configure simulation samples per duration}
  \State $D \gets floor(s / increment)$ \Comment{Durations are sampled at discrete intervals}
  \State $P \gets Array(D+1,D+1)$       \Comment{The distribution P is a 2 dimensional array}
  \For{$d \in 0 to D$}                  \Comment{Iterate over all possible true gaze durations}
    \For{$n \in 0 to N$}                \Comment{Iterature to collect the N samples for d}
      \State $p \gets generatePath(d,l,s)$
      \State $d_e \gets generateMeasurement(p,l,C)$  
      \State $P[d,d_e] += 1/N$          \Comment{Probability increment for a measuremnt of $d_e$}
    \EndFor
  \EndFor
  \State return P
\end{algorithmic}
\end{algorithm}

The bulk of the work is done by two functions inside the simulation loop. The first function
$generatePath(d,l,s)$ takes a true gaze duration (for that simulation iteration) the location
of the gaze target and the viewing session length. It then
generates a random path of fixation points that is compatible with those parameters.
The second function 
$generateMeasurement(p,l,C)$, takes the path, target location and the eye tracking calibration 
data and produces a sample of measured attention time. It does this by using the calibration data to sample
fixation errors and apply them to the path data. The noisy path data is then used to determine
the estimated gaze duration by looking at the number of fixation points that intersect with the target
location.
Repeated application of this process for a given value of $d$ produces multiple samples of $\hat{d}$
for that true gaze duration. Repeating the process over different 
values of $d$ allows us to populate a two dimensional array where the first dimension is the
true duration and the second the estimated duration. Note that our probability distribution is discrete
because we restrict ourselves to sampling over a fixed set of increments 
between zero and the session length.

\subsection{Gaze Duration Distribution}

In real world applications we will have the true session length but rely on the model for the 
estimate of gaze duration. Meaning that the probability distribution we want is the distribution 
over true gaze durations for a given measurement, rather than the inverse which is what we 
have produced in the previous section. We produce an estimate of $P(d|\hat{d},l,s)$
through an application of Bayes' rule, as shown in Equation \ref{eq:p_of_d}.

\begin{equation}
\label{eq:p_of_d}
P(d|\hat{d},l,s) =  \frac{ P(\hat{d},l,s|d) \dot P(d) }{ P(\hat{d},l,s)  }
\end{equation}

We use a uniform prior for $P(d)$, meaning in the absence of additional information all
gaze durations less than the session length are equally likely. As our distributions are
discrete estimations of an underlying continuous distribution the value of $P(d)$ is
equal to $1/(1+floor(10*s))$
 We can estimate the value
of the denominator $P(\hat{d},l,s)$ by iterating over all values of d and summing the product
of $ P(\hat{d},l,s|d) \dot P(d)$. This fully explicated form is shown in 
Equation \ref{eq:p_of_d_full}.

\begin{equation}
\label{eq:p_of_d_full}
P(d|\hat{d},l,s) =  \frac{ P(\hat{d},l,s|d) \dot P(d) }{ \sum_{\delta \in D} P(\hat{d},l,s|\delta) \dot P(\delta)  }
\end{equation}

\subsection{Implementation}

The algorithm described in the preceding sections has been implemented as an open source python 
package called gazerr. It can be used as a software library, or deployed as a command line application. Source
code is available on GitHub () and the package can be installed from PyPi.

The source repository for the gazerr application also contains a series of scripts for generating data
sets and running the experiments outlined in the next section. Full details available in the package
README file.

\subsection{Experiments}

We apply the gazerr application to investigate the relationship between measured gaze duration
and expected real duration under a variety of changing scenarios. These scenarios are variations
in the underlying error indicated by the eye tracking calibration file and length of the session
in which the user could be looking at the area of interest.

For the sake of our simulations we use a fixed sized simulated device with viewport pixel 
dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium rectangle ad unit (MREC)
that is located toward the top of the screen. We iterate over a range of potential mean error
values (determined to be the mean Euclidean distance between the measured fixation and actual fixation).
For each of these mean error values we generate a synthetic eye tracking calibration file.

We feed each of these synthetic calibration files into gazerr to calculate posterior probability
distributions over true gaze duration for the set of measured gaze durations. These posterior
distributions are converted into expected values of true gaze duration for each measured duration.
We repeat the above process for multiple session lengths to investigate the impact of increasing
session length on the expected gaze duration for a fixed set of measurements.

\section{Results}

The first experimental result is the expected true duration against the measured duration
for a single eye tracking application with a mean fixation error of 50 pixels. These
results are shown in Figure \ref{fig:measured_vs_expected}.

\begin{figure}[h!]
\includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
\caption{Measured versus expected gaze duration. MREC in 10s Session with 50px Mean Error.}
\label{fig:measured_vs_expected}
\end{figure}

The result shown in Figure \ref{fig:measured_vs_expected} is indicative of what we see under
multiple variations of the simulation configuration. At lower meaurement values there is a 
tendency for the measurement to be an over estimation, and at higher values an under-estimation.

We next examined the way that exected error in the gaze duration estimate tracks with expected 
error in the gaze fixation points. This involved calculating the mean absolute error in gaze
duration for each of the mean fixation error validation files. We performed this calculation
using two methods. The first method, called uniform, assumes that all measured values are 
equally likely. The second method, called smoothed, assumes that low measurements are much
more likely than higher measurement, and that this drop in likelihood follows an approximately
exponential distribution.

\begin{figure}[h!]
\includegraphics[scale=0.5]{../results/Fixation_vs_duration_error.png}
\caption{Fixation error versus gaze duration error.}
\label{fig:fixation_vs_duration}
\end{figure}

As shown in Figure \ref{fig:fixation_vs_duration} we see that the error in gaze duration
grows with error in fixation points, as would be expected. However, when we use the exponential
weighting to calculate the expected error in duration, the gaze duration error grows much slower
than in the uniform case. This suggests that the observed tendency toward smaller measurements
in area of interest studies means lower expected error.

In our final experiment we look at the effect of both the size of the measured 
gaze duration and the mean error of fixation on the expected error in gaze duration.
We display this as a three dimensional expected error surface over these two key dimensions
in Figure \ref{fig:error_surface}

\begin{figure}[h!]
\includegraphics[scale=0.5]{../results/Error_surface.png}
\caption{Mean gaze duration error over measured error and fixation MAE.}
\label{fig:error_surface}
\end{figure}

There is a region in Figure \ref{fig:error_surface} that is unpopulated due to the fact that
the estimated distributions contained no density in these regions. This was observed for 
eye tracking models with high fixation error and when the measured duration was high. Meaning
that our model estimates that when fixation error is this high, measurements of these values
become extremely unlikley to be observed.

\section{Conclusion}

We have demonstrated that the statistial information contained in eye tracking validation
data can be utilised for estimating errors in gaze duration estimation. The process involves
simulating the distribution of measurements for a given true durastion, and then inverting
it into a distribution over true duration for a given measurement.

We have released this approach as the open source application gazerr, and used it to explore
the relationship between fixation error and duration error. Our results show that the nature
of the error in gaze duration depends on the size of the measurement. Smaller measurements
tend toward over prediction, and larger measurements towards under prediction. We plotted
the relationship between expected error and saw that observed tendency toward smaller duration
measurements (estimated with an psuedo-exponential distribution) delivers a slower growing
expected duraiton error with the mean error in fixation points.

Finally, our simualtions revealed that for large error in fixation points certain measurements
become some unlikely that our monte carlo simualtions placed none of the probability density
in those regions. This strongly suggests that an additional downside to large fixation point error
is a reduced range of measurement for area of interest studies.

\bibliographystyle{IEEEtran}
\bibliography{refs}

\end{document}